### 2017-05-07 paper/main.tex
first commit of paper

\documentclass{article}
\usepackage[utf8]{inputenc}

% Paper layout
\setlength{\paperheight}{11in}
\setlength{\paperwidth}{8.5in}
\oddsidemargin 0.0in 
\evensidemargin .5in
\marginparwidth 0.07 true in
\textheight 7.5 true in
\textwidth 6.5 true in

\PassOptionsToPackage{numbers, compress}{natbib}
\usepackage{url}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage{mathtools}
\usepackage{fancyvrb}
% \usepackage[shortlabels]{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\hmwkTitle}{Extensible Physics Engine}
% \newcommand{\hmwkIssueDate}{22 February 2017}
\newcommand{\hmwkDueDate}{8 May 2017}
\newcommand{\hmwkClass}{6.905/6.945 Final Project Draft}
\newcommand{\hmwkClassInstructor}{Professor Gerry Sussman}
\newcommand{\hmwkAuthorName}{\textbf{Author1, Author2, Author3}}

\title{
    \textmd{\hmwkClass:\ \hmwkTitle}\\
    \\
    % \small{Issued\ on\ \hmwkIssueDate}\\
    \small{Due\ on\ \hmwkDueDate}
}
\author{\hmwkAuthorName}
\date{}
\begin{document}
\maketitle
\section{Introduction}
blah blah blah

\section{What we did}

\section{How we did it}

\section{Why we did it this way}

\section{Demos}

\end{document}

### 2017-05-07 paper/main.tex
more details

@@
 
-\section{What we did}
+\section{Infrastructure}
+generic procedures, ps4 stuff
+what we did, how we did it, why we did it
 
-\section{How we did it}
+\section{Arithmetic}
+vector arithmetic
+what we did, how we did it, why we did it
 
-\section{Why we did it this way}
+\section{Physics Engine}
+each object does its own updating, procedures are objects
+what we did, how we did it, why we did it
 
 \section{Demos}
+binary stars, solar system, magnetics, magnetic solar system
 

### 2017-05-07 paper/main.tex
gravity demos

@@
 \usepackage{fancyvrb}
+\usepackage{array,booktabs,calc,blindtext}
 % \usepackage[shortlabels]{enumerate}
@@
 \section{Demos}
-binary stars, solar system, magnetics, magnetic solar system
+We can easily extend our system with new forces and new objects. We begin by creating an gravity as an \texttt{interaction}:
+
+\begin{verbatim}
+(define (make-gravity thing all-things)
+  (define (procedure thing influences)
+    (sum (map (lambda (influence)
+                (let* ((m1 (get-mass thing))
+                      (m2 (get-mass influence))
+                      (G 6.674e-11)
+                      (v (- (get-position influence)
+                                       (get-position thing)))
+                      (r (magnitude v))
+                      (u (/ v r))
+                      (gmag (* (* G m1 m2)
+                              (/ 1 (square r))))
+                    )
+                  (* u gmag)
+                ))
+              influences)))
+  (let ((influences (delq thing all-things)))
+    (make-interaction gravity? 'gravity procedure influences)))
+\end{verbatim}
+
+Below we've created several worlds that exhibit gravity. Once an object is added to the world, any object that has mass will feel gravitational forces with other object that have mass.
+
+\begin{verbatim}
+(define (earth-moon)
+  (define w (make-world "world"))
+  (define b1 (make-ball "earth" 30 1e15 #(0 0) #(0 0) "blue"))
+  (define b2 (make-ball "moon" 5 1e5 #(100 100) #(-15.361 15.361) "#85929E"))
+  (add-mass! b1 w)
+  (add-mass! b2 w)
+  w)
+\end{verbatim}
+
+\begin{verbatim}
+(define (binary-stars)
+  (define w (make-world "world"))
+  (define b1 (make-ball "ball1" 5 1e15 #(-100 -100) #(9 -9) "red"))
+  (define b2 (make-ball "ball2" 5 1e15 #(100 100) #(-9 9) "green"))
+  (add-mass! b1 w)
+  (add-mass! b2 w)
+  w)
+\end{verbatim}
+
+\begin{verbatim}
+(define (solar-system)
+  (define w (make-world "world"))
+  (define s (make-ball "sun" 30 1e15 #(0 0) #(0 0) "yellow"))
+  (define b1 (make-ball "ball1" 5 1e5 #(100 100) #(-15.361 15.361) "blue"))
+  (define b2 (make-ball "ball2" 5 1e5 #(110 110) #(-15.361 15.361) "red"))
+  (define b3 (make-ball "ball3" 5 1e5 #(120 120) #(-15.361 15.361) "green"))
+  (define b4 (make-ball "ball4" 5 1e5 #(130 130) #(-15.361 15.361) "purple"))
+  (define b5 (make-ball "ball5" 5 1e5 #(140 140) #(-15.361 15.361) "orange"))
+  (define b6 (make-ball "ball6" 5 1e5 #(150 150) #(-15.361 15.361) "gray"))
+
+  (add-mass! s w)
+  (add-mass! b1 w)
+  (add-mass! b2 w)
+  (add-mass! b3 w)
+  (add-mass! b4 w)
+  (add-mass! b5 w)
+  (add-mass! b6 w)
+  w
+)
+\end{verbatim}
+
+\begin{figure}[h!]
+  \centering
+ \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{figs/gravity.png}
+  \caption{\textit{left}: \texttt{earth-moon}; \textit{center}: \texttt{binary-stars}; \textit{right}: \texttt{solar-system}}
+  \label{figure:gravity}
+\end{figure}
+
+\begin{verbatim}
+(define (magnets-1)
+  (define w (make-world "world"))
+  (define m1 (make-magnet "magnet1" 1e5 1 #(-30 -30) #(0 0) "red"))
+  (define m2 (make-magnet "magnet2" 1e5 1 #(30 30) #(0 0) "red"))
+
+  (add-magnet! m1 w)
+  (add-magnet! m2 w)
+  w
+)
+\end{verbatim}
+
+\begin{verbatim}
+(define (magnets-2)
+  (define w (make-world "world"))
+  (define m1 (make-magnet "magnet1" 5e5 1 #(-100 -100) #(0 0) "red"))
+  (define m2 (make-magnet "magnet2" -5e5 1 #(100 100) #(0 0) "blue"))
+  (define m3 (make-magnet "magnet3" 5e5 1 #(-100 100) #(0 0) "red"))
+  (define m4 (make-magnet "magnet4" -5e5 1 #(100 -100) #(0 0) "blue"))
+
+  (add-magnet! m1 w)
+  (add-magnet! m2 w)
+  (add-magnet! m3 w)
+  (add-magnet! m4 w)
+  w
+)
+\end{verbatim}
+
+\begin{verbatim}
+(define (magnetic-solar-system)
+  (define w (make-world "world"))
+  (define s (make-magnet "sun" 5e5 1e15 #(0 0) #(0 0) "blue"))
+  (define b1 (make-magnet "ball1" -2e7 1e5 #(100 100) #(-15.361 15.361) "red"))
+  (define b2 (make-magnet "ball2" -2e7 1e5 #(110 110) #(-15.361 15.361) "red"))
+  (define b3 (make-magnet "ball3" -2e7 1e5 #(120 120) #(-15.361 15.361) "red"))
+  (define b4 (make-magnet "ball4" -2e7 1e5 #(130 130) #(-15.361 15.361) "red"))
+  (define b5 (make-magnet "ball5" -2e7 1e5 #(140 140) #(-15.361 15.361) "red"))
+  (define b6 (make-magnet "ball6" -2e7 1e5 #(150 150) #(-15.361 15.361) "red"))
+
+  (add-magnet! s w)
+  (add-magnet! b1 w)
+  (add-magnet! b2 w)
+  (add-magnet! b3 w)
+  (add-magnet! b4 w)
+  (add-magnet! b5 w)
+  (add-magnet! b6 w)
+  w
+)
+\end{verbatim}
 

### 2017-05-07 paper/main.tex
demos

@@
 
-\begin{verbatim}
+{\small\begin{verbatim}
 (define (make-gravity thing all-things)
@@
                        (G 6.674e-11)
-                      (v (- (get-position influence)
+                       (v (- (get-position influence)  ; vector between influence and thing
                             (get-position thing)))
-                      (r (magnitude v))
-                      (u (/ v r))
-                      (gmag (* (* G m1 m2)
-                              (/ 1 (square r))))
-                    )
-                  (* u gmag)
-                ))
+                       (r (magnitude v))               ; distance between influence and thing
+                       (u (/ v r))                     ; unit vector
+                       (gmag (* (* G m1 m2)            ; magnitude of gravity
+                                (/ 1 (square r)))))
+                  (* u gmag)))                        ; gravitational force vector
               influences)))
@@
     (make-interaction gravity? 'gravity procedure influences)))
-\end{verbatim}
+\end{verbatim}}
 
-Below we've created several worlds that exhibit gravity. Once an object is added to the world, any object that has mass will feel gravitational forces with other object that have mass.
+Below we've created several worlds that exhibit gravity (Fig.
+\ref{figure:gravity}).
+Once an object is added to the world, any object that has mass will feel gravitational forces with other object that have mass.
 
-\begin{verbatim}
+This world contains an earth with a moon orbiting around it.
+{\small\begin{verbatim}
 (define (earth-moon)
@@
   w)
-\end{verbatim}
+\end{verbatim}}
 
-\begin{verbatim}
+This world contains two equal masses orbiting a common center of mass.
+{\small\begin{verbatim}
 (define (binary-stars)
@@
   w)
-\end{verbatim}
+\end{verbatim}}
 
-\begin{verbatim}
+This world contains a massive sun and several planets orbitting around the sun.
+{\small\begin{verbatim}
 (define (solar-system)
@@
   (add-mass! b6 w)
-  w
-)
-\end{verbatim}
+  w)
+\end{verbatim}}
 
@@
  \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{figs/gravity.png}
-  \caption{\textit{left}: \texttt{earth-moon}; \textit{center}: \texttt{binary-stars}; \textit{right}: \texttt{solar-system}}
+  \caption{[\textit{left}] \texttt{earth-moon}; [\textit{center}] \texttt{binary-stars}; [\textit{right}] \texttt{solar-system}}
   \label{figure:gravity}
@@
 
+Let's add magnetism into the physics engine.
+
+\begin{verbatim}
+(define (make-magnetic-force magnet all-magnets)
+  (define (procedure magnet influences)
+    (sum (map (lambda (influence)
+                (let* ((q1 (get-magnet-charge magnet))
+                           (q2 (get-magnet-charge influence))
+                           (mu 1.256e-6)  ; permeability of air
+                           (v (- (get-position influence)  ; vector between influence and magnet
+                             (get-position magnet)))
+                           (r (magnitude v))               ; distance between influence and magnet
+                           (u (* -1 (/ v r)))              ; unit vector
+                           (mmag (* (* mu q1 q2)           ; magnitude of magnetic force
+                                (/ 1 (* 4 pi (square r))))))
+                  (* u mmag)))                             ; magnetic force vector
+              influences)))
+  (let ((influences (delq magnet all-magnets)))
+    (make-interaction magnetic-force? 'magnetic-force procedure influences)))
+\end{verbatim}
+
+This is a simple world that two magnets repelling each other.
 \begin{verbatim}
@@
 
+This world contains two positively charged masses (blue) and two negatively charged masses (red).
 \begin{verbatim}
@@
 \end{verbatim}
+
+We can also combine gravity and magnetism together.
+Here, we created a solar-system-like world where each of the planets repel each other and the sun attracts the planets via gravity and magnetism.
+As figure \ref{figure:magnetism} (right) shows, this magnetic solar system exhibits wider and diverging orbits compared to figure \ref{figure:gravity} (right), the solar system with only gravity.
 
@@
 
+\begin{figure}[h!]
+  \centering
+ \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{figs/magnetism.png}
+  \caption{[\textit{left}] \texttt{magnets-1}; [\textit{center}] \texttt{magnets-2}; [\textit{right}] \texttt{magnetic-solar-system}}
+  \label{figure:magnetism}
+\end{figure}
+
 \end{document}

### 2017-05-07 paper/main.tex
arithmetic

--- 2017-05-07 paper/main.tex
@@
 \newcommand{\hmwkClass}{6.905/6.945 Final Project Draft}
-\newcommand{\hmwkClassInstructor}{Professor Gerry Sussman}
+% \newcommand{\hmwkClassInstructor}{Professor Gerry Sussman}
 \newcommand{\hmwkAuthorName}{\textbf{Author1, Author2, Author3}}
@@
     \textmd{\hmwkClass:\ \hmwkTitle}\\
-    \\
+    \normalsize{\hmwkClassInstructor}\\
     % \small{Issued\ on\ \hmwkIssueDate}\\
@@
 \section{Arithmetic}
-vector arithmetic
-what we did, how we did it, why we did it
+We incorporated the generic arithmetic developed in pset 3.
+We extended the vector-arithmetic to support element-wise operations between scalars and vectors.
+We did this by coercing scalars to become vectors that contain the same value.
+For example, \texttt{(+ \#(3 4) 1)} gets transformed into \texttt{(+ \#(3 4) #(1 1))}.
+The coercian method is
+\begin{verbatim}
+(define (coerce a size-reference)
+  (if (vector? a)
+    a
+    (make-vector (vector-length size-reference) a)))
+\end{verbatim}
+The rest of the infrastructure for generic procedures was adapted from pset 4, but pset 4's generic arithmetic is not compatible with that of pset 3.
+To fix this issue, we first loaded the relevant pset 3 dependencies, installed a generic arithmetic with vector operations, redefined arithmetic operations to be the ones from that generic arithmetic (such as with \texttt{(define * (access * arith-environment))}), and then loaded the pset 4 generic procedure infrastructure on top.
 
@@
 \section{Demos}
-binary stars, solar system, magnetics, magnetic solar system
+\subsection{Gravity}
+% \begin{samepage}
+We can easily extend our system with new forces and new objects. We begin by creating an gravity as an \texttt{interaction}:
+{\small\begin{verbatim}
+(define (make-gravity thing all-things)
+  (define (procedure thing influences)
+    (sum (map (lambda (influence)
+                (let* ((m1 (get-mass thing))
+                       (m2 (get-mass influence))
+                       (G 6.674e-11)
+                       (v (- (get-position influence)  ; vector between influence and thing
+                            (get-position thing)))
+                       (r (magnitude v))               ; distance between influence and thing
+                       (u (/ v r))                     ; unit vector
+                       (gmag (* (* G m1 m2)            ; magnitude of gravity
+                                (/ 1 (square r)))))
+                  (* u gmag)))                        ; gravitational force vector
+              influences)))
+  (let ((influences (delq thing all-things)))
+    (make-interaction gravity? 'gravity procedure influences)))
+\end{verbatim}}
+% \end{samepage}
+
+\begin{figure}[h!]
+  \centering
+ \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{figs/gravity.png}
+  \caption{[\textit{left}] \texttt{earth-moon}; [\textit{center}] \texttt{binary-stars}; [\textit{right}] \texttt{solar-system}}
+  \label{figure:gravity}
+\end{figure}
+
+We've created several worlds that exhibit gravity (Fig.
+\ref{figure:gravity}).
+Once an object is added to the world, any object that has mass will feel gravitational forces with other object that have mass.
+
+% \begin{samepage}
+This world contains an earth with a moon orbiting around it.
+{\small\begin{verbatim}
+(define (earth-moon)
+  (define w (make-world "world"))
+  (define b1 (make-ball "earth" 30 1e15 #(0 0) #(0 0) "blue"))
+  (define b2 (make-ball "moon" 5 1e5 #(100 100) #(-15.361 15.361) "#85929E"))
+
+  (add-mass! b1 w)
+  (add-mass! b2 w)
+  w)
+\end{verbatim}}
+% \end{samepage}
+
+% \begin{samepage}
+This world contains two equal masses orbiting a common center of mass.
+{\small\begin{verbatim}
+(define (binary-stars)
+  (define w (make-world "world"))
+  (define b1 (make-ball "ball1" 5 1e15 #(-100 -100) #(9 -9) "red"))
+  (define b2 (make-ball "ball2" 5 1e15 #(100 100) #(-9 9) "green"))
+
+  (add-mass! b1 w)
+  (add-mass! b2 w)
+  w)
+\end{verbatim}}
+% \end{samepage}
+
+% \begin{samepage}
+This world contains a massive sun and several planets orbitting around the sun.
+{\small\begin{verbatim}
+(define (solar-system)
+  (define w (make-world "world"))
+  (define s (make-ball "sun" 30 1e15 #(0 0) #(0 0) "yellow"))
+  (define b1 (make-ball "ball1" 5 1e5 #(100 100) #(-15.361 15.361) "blue"))
+  (define b2 (make-ball "ball2" 5 1e5 #(110 110) #(-15.361 15.361) "red"))
+  (define b3 (make-ball "ball3" 5 1e5 #(120 120) #(-15.361 15.361) "green"))
+  (define b4 (make-ball "ball4" 5 1e5 #(130 130) #(-15.361 15.361) "purple"))
+  (define b5 (make-ball "ball5" 5 1e5 #(140 140) #(-15.361 15.361) "orange"))
+  (define b6 (make-ball "ball6" 5 1e5 #(150 150) #(-15.361 15.361) "gray"))
+
+  (add-mass! s w)
+  (add-mass! b1 w)
+  (add-mass! b2 w)
+  (add-mass! b3 w)
+  (add-mass! b4 w)
+  (add-mass! b5 w)
+  (add-mass! b6 w)
+  w)
+\end{verbatim}}
+% \end{samepage}
+
+\subsection{Magentism}
+% \begin{samepage}
+Let's add magnetism into the physics engine. 
+{\small\begin{verbatim}
+(define (make-magnetic-force magnet all-magnets)
+  (define (procedure magnet influences)
+    (sum (map (lambda (influence)
+                (let* ((q1 (get-magnet-charge magnet))
+                           (q2 (get-magnet-charge influence))
+                           (mu 1.256e-6)  ; permeability of air
+                           (v (- (get-position influence)  ; vector between influence and magnet
+                             (get-position magnet)))
+                           (r (magnitude v))               ; distance between influence and magnet
+                           (u (* -1 (/ v r)))              ; unit vector
+                           (mmag (* (* mu q1 q2)           ; magnitude of magnetic force
+                                (/ 1 (* 4 pi (square r))))))
+                  (* u mmag)))                             ; magnetic force vector
+              influences)))
+  (let ((influences (delq magnet all-magnets)))
+    (make-interaction magnetic-force? 'magnetic-force procedure influences)))
+\end{verbatim}}
+% \end{samepage}
+
+% \begin{samepage}
+This is a simple world that two magnets repelling each other.
+{\small\begin{verbatim}
+(define (magnets-1)
+  (define w (make-world "world"))
+  (define m1 (make-magnet "magnet1" 1e5 1 #(-30 -30) #(0 0) "red"))
+  (define m2 (make-magnet "magnet2" 1e5 1 #(30 30) #(0 0) "red"))
+
+  (add-magnet! m1 w)
+  (add-magnet! m2 w)
+  w)
+\end{verbatim}}
+% \end{samepage}
+
+% \begin{samepage}
+This world contains two positively charged masses (blue) and two negatively charged masses (red).
+{\small\begin{verbatim}
+(define (magnets-2)
+  (define w (make-world "world"))
+  (define m1 (make-magnet "magnet1" 5e5 1 #(-100 -100) #(0 0) "red"))
+  (define m2 (make-magnet "magnet2" -5e5 1 #(100 100) #(0 0) "blue"))
+  (define m3 (make-magnet "magnet3" 5e5 1 #(-100 100) #(0 0) "red"))
+  (define m4 (make-magnet "magnet4" -5e5 1 #(100 -100) #(0 0) "blue"))
+
+  (add-magnet! m1 w)
+  (add-magnet! m2 w)
+  (add-magnet! m3 w)
+  (add-magnet! m4 w)
+  w)
+\end{verbatim}}
+% \end{samepage}
+
+% \begin{samepage}
+We can also combine gravity and magnetism together.
+Here, we created a solar-system-like world where each of the planets repel each other and the sun attracts the planets via gravity and magnetism.
+As figure \ref{figure:magnetism} (right) shows, this magnetic solar system exhibits wider and diverging orbits compared to figure \ref{figure:gravity} (right), the solar system with only gravity.
+{\small\begin{verbatim}
+(define (magnetic-solar-system)
+  (define w (make-world "world"))
+  (define s (make-magnet "sun" 5e5 1e15 #(0 0) #(0 0) "blue"))
+  (define b1 (make-magnet "ball1" -2e7 1e5 #(100 100) #(-15.361 15.361) "red"))
+  (define b2 (make-magnet "ball2" -2e7 1e5 #(110 110) #(-15.361 15.361) "red"))
+  (define b3 (make-magnet "ball3" -2e7 1e5 #(120 120) #(-15.361 15.361) "red"))
+  (define b4 (make-magnet "ball4" -2e7 1e5 #(130 130) #(-15.361 15.361) "red"))
+  (define b5 (make-magnet "ball5" -2e7 1e5 #(140 140) #(-15.361 15.361) "red"))
+  (define b6 (make-magnet "ball6" -2e7 1e5 #(150 150) #(-15.361 15.361) "red"))
+
+  (add-magnet! s w)
+  (add-magnet! b1 w)
+  (add-magnet! b2 w)
+  (add-magnet! b3 w)
+  (add-magnet! b4 w)
+  (add-magnet! b5 w)
+  (add-magnet! b6 w)
+  w)
+\end{verbatim}}
+% \end{samepage}
+
+\begin{figure}[h!]
+  \centering
+ \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{figs/magnetism.png}
+  \caption{[\textit{left}] \texttt{magnets-1}; [\textit{center}] \texttt{magnets-2}; [\textit{right}] \texttt{magnetic-solar-system}}
+  \label{figure:magnetism}
+\end{figure}
 

### 2017-05-07 paper/main.tex
Added stuffs under intro and physics engine

@@
 \section{Introduction}
-blah blah blah
+motivation: features of physics engine that incorporate ideas from class (e.g. generic operator, hierarchy, propagator, language for language (high--low)), cool visualization of concepts from class
 
@@
 We did this by coercing scalars to become vectors that contain the same value.
-For example, \texttt{(+ \#(3 4) 1)} gets transformed into \texttt{(+ \#(3 4) #(1 1))}.
-The coercian method is
+For example, \texttt{(+ \#(3 4) 1)} gets transformed into \texttt{(+ \#(3 4) \#(1 1))}.
+
+This extension was made to allow our system to operate on any $n$th dimension vector, and therefore be applicable to any $n$th dimension world representation.
+The default velocity of an object is set to 0 if it is not specified.
+Supporting arithmetic operations between a scalar and a vector enables our system to use a single default value 0 that is compatible with any vector instead of limiting our system to the dimension of the default zero vector specified.
+
+The coercion method is 
 \begin{verbatim}
@@
 \section{Physics Engine}
-each object does its own updating, procedures are objects
+There are mainly two ways we could update object positions: 1) The evolver can take in a list of forces and objects and apply forces as appropriate 2) Each object can be registered with a set of forces applicable to them and the evolver can call each object to evolve themselves.
+
+We chose the second model where each object does its own updating because [blank].
+
+There are three types of objects in the system: $thing$, $interaction$, and $world$.
+$thing$ refers to all physical objects in the system.
+There is a hierarchy of physical objects that have additional properties.
+For example, type $magnet$ has an additional property $charge$ and experiences magnetic forces while type $ball$ experiences only gravity.
+Each $thing$ keeps a list of $interactions$ that is used to update itself.
+$interaction$ refers to forces exerted on $thing$.
+$interaction$ includes a procedure that calculates the force as well as $influence$, a list of $thing$ that contribute to that force.
+Similar to $thing$, subtypes of $interaction$  form a hierarchy.
+Lastly, $world$ keeps a list of all $things$ and an $update$ call to the $world$ calls on each $thing$ to update itself.
+
 what we did, how we did it, why we did it
+
+(subject to change after restructuring)
 

### 2017-05-07 paper/main.tex
physics engine

--- 2017-05-07 paper/main.tex
@@
 what we did, how we did it, why we did it
+
+Our engine updates each object every timestep and re-renders them on the screen:
+\begin{verbatim}
+(define (run-engine world steps)
+  (reset-graphics)
+  (if (> steps 0)
+    (begin 
+      (for-each 
+        (lambda (thing)
+            (newline)
+            (display (cons (get-name thing) (get-position thing)))  ; for debugging
+            (render thing)
+          )
+          (get-world-all-things world))
+      (update-world world)
+      (run-engine world (- steps 1)))))
+\end{verbatim}
+
+where the \texttt{render} procedure takes in a \texttt{thing}, checks what type of thing it is, and renders the object appropriately.
 

### 2017-05-07 paper/main.tex
formatting

--- 2017-05-07 paper/main.tex
@@
 what we did, how we did it, why we did it
+
+Our engine updates each object every timestep and re-renders them on the screen:
+\begin{verbatim}
+(define (run-engine world steps)
+  (reset-graphics)
+  (if (> steps 0)
+    (begin 
+      (for-each 
+        (lambda (thing)
+            (newline)
+            (display (cons (get-name thing) (get-position thing)))  ; for debugging
+            (render thing))
+        (get-world-all-things world))
+      (update-world world)
+    (run-engine world (- steps 1)))))
+\end{verbatim}
+
+where the \texttt{render} procedure takes in a \texttt{thing}, checks what type of thing it is, and renders the object appropriately.
 

### 2017-05-07 paper/main.tex

--- 2017-05-07 paper/main.tex
@@
 (subject to change after restructuring)
+
+Our engine updates each object every timestep and re-renders them on the screen:
+\begin{verbatim}
+(define (run-engine world steps)
+  (reset-graphics)
+  (if (> steps 0)
+    (begin 
+      (for-each 
+        (lambda (thing)
+            (newline)
+            (display (cons (get-name thing) (get-position thing)))  ; for debugging
+            (render thing)
+          )
+          (get-world-all-things world))
+      (update-world world)
+      (run-engine world (- steps 1)))))
+\end{verbatim}
+
+where the \texttt{render} procedure takes in a \texttt{thing}, checks what type of thing it is, and renders the object appropriately.
 

### 2017-05-07 paper/main.tex

@@
             (display (cons (get-name thing) (get-position thing)))  ; for debugging
-            (render thing)
-          )
+            (render thing))
         (get-world-all-things world))

### 2017-05-08 paper/main.tex
Added to physics engine

@@
 \section{Physics Engine}
-There are mainly two ways we could update object positions: 1) The evolver can take in a list of forces and objects and apply forces as appropriate 2) Each object can be registered with a set of forces applicable to them and the evolver can call each object to evolve themselves.
-
-We chose the second model where each object does its own updating because [blank].
-
-There are three types of objects in the system: $thing$, $interaction$, and $world$.
-$thing$ refers to all physical objects in the system.
+\subsection{Object Types}
+There are three types of objects in the system: \texttt{thing}, \texttt{interaction}, and \texttt{world}.
+\texttt{thing} refers to all physical objects in the system.
 There is a hierarchy of physical objects that have additional properties.
-For example, type $magnet$ has an additional property $charge$ and experiences magnetic forces while type $ball$ experiences only gravity.
-Each $thing$ keeps a list of $interactions$ that is used to update itself.
-$interaction$ refers to forces exerted on $thing$.
-$interaction$ includes a procedure that calculates the force as well as $influence$, a list of $thing$ that contribute to that force.
-Similar to $thing$, subtypes of $interaction$  form a hierarchy.
-Lastly, $world$ keeps a list of all $things$ and an $update$ call to the $world$ calls on each $thing$ to update itself.
+For example, type \texttt{magnet} has an additional property \texttt{charge} and experiences magnetic forces while type \texttt{ball} experiences only gravity.
+Each \texttt{thing} keeps a list of \texttt{interaction} that is used to update itself.
+\texttt{interaction} refers to forces exerted on \texttt{thing}s.
+\texttt{interaction} includes a procedure that calculates the force as well as \texttt{influences}, a list of \texttt{thing}s that contribute to that force.
+Similar to \texttt{thing}, subtypes of \texttt{interaction}s form a hierarchy.
+Lastly, \texttt{world} keeps a list of all \texttt{thing}s and an \texttt{update} call to the \texttt{world} calls on each \texttt{thing} to update itself.
 
@@
 
+\subsection{Updating Physical Objects}
+There are mainly two ways we could update object positions: 1) The evolver can take in a list of forces and objects and apply forces as appropriate 2) Each object can be registered with a set of forces applicable to them and the evolver can call each object to evolve themselves.
+
+We chose the second model where each object does its own updating because [blank].
+
+\subsection{Output Display}
 Our engine updates each object every timestep and re-renders them on the screen:

### 2017-05-08 paper/main.tex
Added reason for objects updating itself

@@
 
-what we did, how we did it, why we did it
-
 (subject to change after restructuring)
@@
 \subsection{Updating Physical Objects}
-There are mainly two ways we could update object positions: 1) The evolver can take in a list of forces and objects and apply forces as appropriate 2) Each object can be registered with a set of forces applicable to them and the evolver can call each object to evolve themselves.
+There are mainly two ways positions of physical objects could be updated: 1) The evolver can take in a list of forces and objects and apply forces as appropriate 2) Each object can be registered with a set of forces applicable to it and the evolver can call each object to evolve itself.
 
-We chose the second model where each object does its own updating because [blank].
+Our system adopted the second model where each object updates itself.
+This model makes it easier to handle different types of forces exerted on specific types of object.
+Moreover, in the second model, the engine does not need to be notified and updated everytime a new object is added.
+Instead, update or addition of new \texttt{interaction}s is taken care of at the object level when it is added.
 

### 2017-05-08 paper/main.tex
Added section for future extensions
changed magnetic charges to electric charges

@@
     \textmd{\hmwkClass:\ \hmwkTitle}\\
-    \normalsize{\hmwkClassInstructor}\\
+    % \normalsize{\hmwkClassInstructor}\\
     % \small{Issued\ on\ \hmwkIssueDate}\\
@@
 There is a hierarchy of physical objects that have additional properties.
-For example, type \texttt{magnet} has an additional property \texttt{charge} and experiences magnetic forces while type \texttt{ball} experiences only gravity.
+For example, type \texttt{point-charge} has an additional property \texttt{charge} and experiences electric forces while type \texttt{ball} experiences only gravity.
 Each \texttt{thing} keeps a list of \texttt{interaction} that is used to update itself.
@@
 Instead, update or addition of new \texttt{interaction}s is taken care of at the object level when it is added.
+
+(TODO: add eval/apply code after restructuring)
 
@@
 % \begin{samepage}
-Let's add magnetism into the physics engine. 
+Let's add electric forces between point charges into the physics engine.
 {\small\begin{verbatim}
-(define (make-magnetic-force magnet all-magnets)
-  (define (procedure magnet influences)
+(define (make-electric-force point-charge all-point-charges)
+  (define (procedure point-charge influences)
     (sum (map (lambda (influence)
-                (let* ((q1 (get-magnet-charge magnet))
-                           (q2 (get-magnet-charge influence))
-                           (mu 1.256e-6)  ; permeability of air
-                           (v (- (get-position influence)  ; vector between influence and magnet
-                             (get-position magnet)))
-                           (r (magnitude v))               ; distance between influence and magnet
-                           (u (* -1 (/ v r)))              ; unit vector
-                           (mmag (* (* mu q1 q2)           ; magnitude of magnetic force
-                                (/ 1 (* 4 pi (square r))))))
-                  (* u mmag)))                             ; magnetic force vector
+                (let* ((q1 (get-point-charge-charge point-charge))
+            		       (q2 (get-point-charge-charge influence))
+            		       (epsilon0 8.854e-12) ; vacuum permittivity
+                       ;; vector between influence and point charge
+            		       (v (- (get-position influence)
+                             (get-position point-charge)))
+                       ;; distance between influence and point charge
+            		       (r (magnitude v))
+                       ;; unit vector * -1 to account for opposite signs
+                       ;; attract & same signs repel
+            		       (u (* -1 (/ v r)))
+            		       (melec (/ (* q1 q2)  ; magnitude of electric force
+                                 (* 4 pi epsilon0 (square r)))))
+                  (* u melec))) ; multiply unit vector by magnitude of force
               influences)))
-  (let ((influences (delq magnet all-magnets)))
-    (make-interaction magnetic-force? 'magnetic-force procedure influences)))
+  (let ((influences (delq point-charge all-point-charges)))
+    (make-interaction electric-force? 'electric-force procedure influences)))
 \end{verbatim}}
@@
 % \begin{samepage}
-This is a simple world that two magnets repelling each other.
+This is a simple world with two point charges repelling each other.
 {\small\begin{verbatim}
-(define (magnets-1)
+(define (charges-1)
   (define w (make-world "world"))
-  (define m1 (make-magnet "magnet1" 1e5 1 #(-30 -30) #(0 0) "red"))
-  (define m2 (make-magnet "magnet2" 1e5 1 #(30 30) #(0 0) "red"))
+  (define m1 (make-point-charge "charge1" 1e-3 1 #(-30 -30) #(0 0) "red"))
+  (define m2 (make-point-charge "charge2" 1e-3 1 #(30 30) #(0 0) "red"))
 
-  (add-magnet! m1 w)
-  (add-magnet! m2 w)
+  (add-point-charge! m1 w)
+  (add-point-charge! m2 w)
   w)
@@
 {\small\begin{verbatim}
-(define (magnets-2)
+(define (charges-2)
   (define w (make-world "world"))
-  (define m1 (make-magnet "magnet1" 5e5 1 #(-100 -100) #(0 0) "red"))
-  (define m2 (make-magnet "magnet2" -5e5 1 #(100 100) #(0 0) "blue"))
-  (define m3 (make-magnet "magnet3" 5e5 1 #(-100 100) #(0 0) "red"))
-  (define m4 (make-magnet "magnet4" -5e5 1 #(100 -100) #(0 0) "blue"))
+  (define m1 (make-point-charge "charge1" 1e-3 1 #(-100 -100) #(0 0) "red"))
+  (define m2 (make-point-charge "charge2" -1e-3 1 #(100 100) #(0 0) "blue"))
+  (define m3 (make-point-charge "charge3" 1e-3 1 #(-100 100) #(0 0) "red"))
+  (define m4 (make-point-charge "charge4" -1e-3 1 #(100 -100) #(0 0) "blue"))
 
-  (add-magnet! m1 w)
-  (add-magnet! m2 w)
-  (add-magnet! m3 w)
-  (add-magnet! m4 w)
+  (add-point-charge! m1 w)
+  (add-point-charge! m2 w)
+  (add-point-charge! m3 w)
+  (add-point-charge! m4 w)
   w)
@@
 % \begin{samepage}
-We can also combine gravity and magnetism together.
-Here, we created a solar-system-like world where each of the planets repel each other and the sun attracts the planets via gravity and magnetism.
-As figure \ref{figure:magnetism} (right) shows, this magnetic solar system exhibits wider and diverging orbits compared to figure \ref{figure:gravity} (right), the solar system with only gravity.
+We can also combine gravity and electric charge.
+Here, we created a solar-system-like world where each of the planets repel each other and the sun attracts the planets via gravity and electric force.
+As figure \ref{figure:charges} (right) shows, this charged solar system exhibits wider and diverging orbits compared to figure \ref{figure:gravity} (right), the solar system with only gravity.
 {\small\begin{verbatim}
-(define (magnetic-solar-system)
+(define (charged-solar-system)
   (define w (make-world "world"))
-  (define s (make-magnet "sun" 5e5 1e15 #(0 0) #(0 0) "blue"))
-  (define b1 (make-magnet "ball1" -2e7 1e5 #(100 100) #(-15.361 15.361) "red"))
-  (define b2 (make-magnet "ball2" -2e7 1e5 #(110 110) #(-15.361 15.361) "red"))
-  (define b3 (make-magnet "ball3" -2e7 1e5 #(120 120) #(-15.361 15.361) "red"))
-  (define b4 (make-magnet "ball4" -2e7 1e5 #(130 130) #(-15.361 15.361) "red"))
-  (define b5 (make-magnet "ball5" -2e7 1e5 #(140 140) #(-15.361 15.361) "red"))
-  (define b6 (make-magnet "ball6" -2e7 1e5 #(150 150) #(-15.361 15.361) "red"))
+  (define s (make-point-charge "sun" 3e-3 1e15 #(0 0) #(0 0) "blue"))
+  (define b1 (make-point-charge "ball1" -1e-1 1e5 #(100 100) #(-15.361 15.361) "red"))
+  (define b2 (make-point-charge "ball2" -1e-1 1e5 #(110 110) #(-15.361 15.361) "red"))
+  (define b3 (make-point-charge "ball3" -1e-1 1e5 #(120 120) #(-15.361 15.361) "red"))
+  (define b4 (make-point-charge "ball4" -1e-1 1e5 #(130 130) #(-15.361 15.361) "red"))
+  (define b5 (make-point-charge "ball5" -1e-1 1e5 #(140 140) #(-15.361 15.361) "red"))
+  (define b6 (make-point-charge "ball6" -1e-1 1e5 #(150 150) #(-15.361 15.361) "red"))
 
-  (add-magnet! s w)
-  (add-magnet! b1 w)
-  (add-magnet! b2 w)
-  (add-magnet! b3 w)
-  (add-magnet! b4 w)
-  (add-magnet! b5 w)
-  (add-magnet! b6 w)
-  w)
 \end{verbatim}}
@@
   \centering
- \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{figs/magnetism.png}
-  \caption{[\textit{left}] \texttt{magnets-1}; [\textit{center}] \texttt{magnets-2}; [\textit{right}] \texttt{magnetic-solar-system}}
-  \label{figure:magnetism}
+ \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{figs/charges.png}
+  \caption{[\textit{left}] \texttt{charges-1}; [\textit{center}] \texttt{charges-2}; [\textit{right}] \texttt{charged-solar-system}}
+  \label{figure:charges}
 \end{figure}
 
+\section{Extensibility of the System}
+Possible future extensions and high level explanation of how they would be implemented (through minimal change to current system)
+
 \end{document}

### 2017-05-08 paper/main.tex
changed to electromagnetism

--- 2017-05-08 paper/main.tex
@@
 \section{Introduction}
-motivation: features of physics engine that incorporate ideas from class (e.g. generic operator, hierarchy, propagator, language for language (high--low)), cool visualization of concepts from class
+
+motivation: features of physics engine that incorporate ideas from class (e.g.
+generic operator, hierarchy, propagator, language for language (high--low)),
+cool visualization of concepts from class
 
@@
 \section{Arithmetic}
-We incorporated the generic arithmetic developed in pset 3.
-We extended the vector-arithmetic to support element-wise operations between scalars and vectors.
-We did this by coercing scalars to become vectors that contain the same value.
-For example, \texttt{(+ \#(3 4) 1)} gets transformed into \texttt{(+ \#(3 4) \#(1 1))}.
 
-This extension was made to allow our system to operate on any $n$th dimension vector, and therefore be applicable to any $n$th dimension world representation.
-The default velocity of an object is set to 0 if it is not specified.
-Supporting arithmetic operations between a scalar and a vector enables our system to use a single default value 0 that is compatible with any vector instead of limiting our system to the dimension of the default zero vector specified.
+We incorporated the generic arithmetic developed in pset 3. We extended the
+vector-arithmetic to support element-wise operations between scalars and
+vectors. We did this by coercing scalars to become vectors that contain the
+same value. For example, \texttt{(+ \#(3 4) 1)} gets transformed into
+\texttt{(+ \#(3 4) \#(1 1))}.
+
+This extension was made to allow our system to operate on any $n$th dimension
+vector, and therefore be applicable to any $n$th dimension world
+representation. The default velocity of an object is set to 0 if it is not
+specified. Supporting arithmetic operations between a scalar and a vector
+enables our system to use a single default value 0 that is compatible with any
+vector instead of limiting our system to the dimension of the default zero
+vector specified.
 
@@
 \end{verbatim}
-The rest of the infrastructure for generic procedures was adapted from pset 4, but pset 4's generic arithmetic is not compatible with that of pset 3.
-To fix this issue, we first loaded the relevant pset 3 dependencies, installed a generic arithmetic with vector operations, redefined arithmetic operations to be the ones from that generic arithmetic (such as with \texttt{(define * (access * arith-environment))}), and then loaded the pset 4 generic procedure infrastructure on top.
+
+The rest of the infrastructure for generic procedures was adapted from pset 4,
+but pset 4's generic arithmetic is not compatible with that of pset 3. To fix
+this issue, we first loaded the relevant pset 3 dependencies, installed a
+generic arithmetic with vector operations, redefined arithmetic operations to
+be the ones from that generic arithmetic (such as with \texttt{(define *
+(access * arith-environment))}), and then loaded the pset 4 generic procedure
+infrastructure on top.
 
@@
 \subsection{Object Types}
-There are three types of objects in the system: \texttt{thing}, \texttt{interaction}, and \texttt{world}.
-\texttt{thing} refers to all physical objects in the system.
-There is a hierarchy of physical objects that have additional properties.
-For example, type \texttt{magnet} has an additional property \texttt{charge} and experiences magnetic forces while type \texttt{ball} experiences only gravity.
-Each \texttt{thing} keeps a list of \texttt{interaction} that is used to update itself.
-\texttt{interaction} refers to forces exerted on \texttt{thing}s.
-\texttt{interaction} includes a procedure that calculates the force as well as \texttt{influences}, a list of \texttt{thing}s that contribute to that force.
-Similar to \texttt{thing}, subtypes of \texttt{interaction}s form a hierarchy.
-Lastly, \texttt{world} keeps a list of all \texttt{thing}s and an \texttt{update} call to the \texttt{world} calls on each \texttt{thing} to update itself.
+
+There are three types of objects in the system: \texttt{thing},
+\texttt{interaction}, and \texttt{world}. \texttt{thing} refers to all physical
+objects in the system. There is a hierarchy of physical objects that have
+additional properties. For example, type \texttt{magnet} has an additional
+property \texttt{charge} and experiences magnetic forces while type
+\texttt{ball} experiences only gravity. Each \texttt{thing} keeps a list of
+\texttt{interaction} that is used to update itself. \texttt{interaction} refers
+to forces exerted on \texttt{thing}s. \texttt{interaction} includes a procedure
+that calculates the force as well as \texttt{influences}, a list of
+\texttt{thing}s that contribute to that force. Similar to \texttt{thing},
+subtypes of \texttt{interaction}s form a hierarchy. Lastly, \texttt{world}
+keeps a list of all \texttt{thing}s and an \texttt{update} call to the
+\texttt{world} calls on each \texttt{thing} to update itself.
 
@@
 \subsection{Updating Physical Objects}
-There are mainly two ways positions of physical objects could be updated: 1) The evolver can take in a list of forces and objects and apply forces as appropriate 2) Each object can be registered with a set of forces applicable to it and the evolver can call each object to evolve itself.
 
-Our system adopted the second model where each object updates itself.
-This model makes it easier to handle different types of forces exerted on specific types of object.
-Moreover, in the second model, the engine does not need to be notified and updated everytime a new object is added.
-Instead, update or addition of new \texttt{interaction}s is taken care of at the object level when it is added.
+There are mainly two ways positions of physical objects could be updated: 1)
+The evolver can take in a list of forces and objects and apply forces as
+appropriate 2) Each object can be registered with a set of forces applicable to
+it and the evolver can call each object to evolve itself.
+
+Our system adopted the second model where each object updates itself. This
+model makes it easier to handle different types of forces exerted on specific
+types of object. Moreover, in the second model, the engine does not need to be
+notified and updated everytime a new object is added. Instead, update or
+addition of new \texttt{interaction}s is taken care of at the object level when
+it is added.
+
+(TODO: add eval/apply code after restructuring)
 
 \subsection{Output Display}
-Our engine updates each object every timestep and re-renders them on the screen:
+
+Our engine updates each object every timestep and re-renders them on the
+screen:
+
 \begin{verbatim}
@@
 
-where the \texttt{render} procedure takes in a \texttt{thing}, checks what type of thing it is, and renders the object appropriately.
+where the \texttt{render} procedure takes in a \texttt{thing}, checks what type
+of thing it is, and renders the object appropriately.
 
@@
 % \begin{samepage}
-We can easily extend our system with new forces and new objects. We begin by creating an gravity as an \texttt{interaction}:
+
+We can easily extend our system with new forces and new objects. We begin by
+creating an gravity as an \texttt{interaction}:
+
 {\small\begin{verbatim}
@@
 
-We've created several worlds that exhibit gravity (Fig.
-\ref{figure:gravity}).
-Once an object is added to the world, any object that has mass will feel gravitational forces with other object that have mass.
+We've created several worlds that exhibit gravity (Fig. \ref{figure:gravity}).
+Once an object is added to the world, any object that has mass will feel
+gravitational forces with other object that have mass.
 
@@
 % \begin{samepage}
+
 This world contains two equal masses orbiting a common center of mass.
+
 {\small\begin{verbatim}
@@
 % \begin{samepage}
+
 This world contains a massive sun and several planets orbitting around the sun.
+
 {\small\begin{verbatim}
@@
 
-\subsection{Magentism}
+\subsection{Electromagentism}
 % \begin{samepage}
-Let's add magnetism into the physics engine. 
+
+Let's add electromagnetism into the physics engine.
+
 {\small\begin{verbatim}
@@
 % \begin{samepage}
-This is a simple world that two magnets repelling each other.
+
+This is a simple world with two charges repelling each other.
+
 {\small\begin{verbatim}
@@
 % \begin{samepage}
-This world contains two positively charged masses (blue) and two negatively charged masses (red).
+
+This world contains two positively charged masses (blue) and two negatively
+charged masses (red).
+
 {\small\begin{verbatim}
@@
 % \begin{samepage}
-We can also combine gravity and magnetism together.
-Here, we created a solar-system-like world where each of the planets repel each other and the sun attracts the planets via gravity and magnetism.
-As figure \ref{figure:magnetism} (right) shows, this magnetic solar system exhibits wider and diverging orbits compared to figure \ref{figure:gravity} (right), the solar system with only gravity.
+
+We can also combine gravity and electromagnetism together. Here, we created a
+solar-system-like world where each of the planets repel each other and the sun
+attracts the planets via gravity and electromagnetism. As figure
+\ref{figure:electromagnetism} (right) shows, this electromagnetism solar system
+exhibits wider and diverging orbits compared to figure \ref{figure:gravity}
+(right), the solar system with only gravity.
+
 {\small\begin{verbatim}
@@
   \caption{[\textit{left}] \texttt{magnets-1}; [\textit{center}] \texttt{magnets-2}; [\textit{right}] \texttt{magnetic-solar-system}}
-  \label{figure:magnetism}
+  \label{figure:electromagnetism}
 \end{figure}
 
+\section{Extensibility of the System}
+
+Possible future extensions and high level explanation of how they would be
+implemented (through minimal change to current system)
+
 \end{document}

### 2017-05-08 paper/main.tex
fixed merge conflict

--- 2017-05-08 paper/main.tex
@@
 \subsection{Object Types}
+
 There are three types of objects in the system: \texttt{thing}, \texttt{interaction}, and \texttt{world}.
@@
 \subsection{Updating Physical Objects}
-There are mainly two ways positions of physical objects could be updated: 1) The evolver can take in a list of forces and objects and apply forces as appropriate 2) Each object can be registered with a set of forces applicable to it and the evolver can call each object to evolve itself.
 
-Our system adopted the second model where each object updates itself.
-This model makes it easier to handle different types of forces exerted on specific types of object.
-Moreover, in the second model, the engine does not need to be notified and updated everytime a new object is added.
-Instead, update or addition of new \texttt{interaction}s is taken care of at the object level when it is added.
+There are mainly two ways positions of physical objects could be updated: 1)
+The evolver can take in a list of forces and objects and apply forces as
+appropriate 2) Each object can be registered with a set of forces applicable to
+it and the evolver can call each object to evolve itself.
+
+Our system adopted the second model where each object updates itself. This
+model makes it easier to handle different types of forces exerted on specific
+types of object. Moreover, in the second model, the engine does not need to be
+notified and updated everytime a new object is added. Instead, update or
+addition of new \texttt{interaction}s is taken care of at the object level when
+it is added.
 
@@
 \subsection{Output Display}
-Our engine updates each object every timestep and re-renders them on the screen:
+
+Our engine updates each object every timestep and re-renders them on the
+screen:
+
 \begin{verbatim}
@@
 
-\subsection{Magentism}
+\subsection{Electromagentism}
 % \begin{samepage}
+
 Let's add electric forces between point charges into the physics engine.
+
 {\small\begin{verbatim}
@@
 % \begin{samepage}
+
 This is a simple world with two point charges repelling each other.
+
 {\small\begin{verbatim}
@@
 % \begin{samepage}
+
 We can also combine gravity and electric charge.
@@
 As figure \ref{figure:charges} (right) shows, this charged solar system exhibits wider and diverging orbits compared to figure \ref{figure:gravity} (right), the solar system with only gravity.
+
 {\small\begin{verbatim}
@@
 \section{Extensibility of the System}
-Possible future extensions and high level explanation of how they would be implemented (through minimal change to current system)
+
+Possible future extensions and high level explanation of how they would be
+implemented (through minimal change to current system)
 

### 2017-05-08 paper/main.tex
added some more content to draft report

--- 2017-05-08 paper/main.tex
@@
 \section{Infrastructure}
-generic procedures, ps4 stuff
+We used the object system from pset 4 to model physical objects in our simulated world, as well as the world itself.
+Having the ability to create a hierarchy of types allows us to derive new types of physical objects to be simulated from previously defined types (see section Object Types).
+
+internal NOTE: generic procedures--we haven't actually used them in the code yet!
+
+NOTE: feel like this should be merged with object types and arithmetic
+
 what we did, how we did it, why we did it
@@
 \section{Extensibility of the System}
-Possible future extensions and high level explanation of how they would be implemented (through minimal change to current system)
+Future extensions include collisions between objects, constraints such as joints, simulating objects in three dimensions, and being able to combine objects into larger systems of objects.
+Collisions and other kinds of constraints could be implemented by adding an interaction that produces forces that enforce these constraints.
+Objects could be simulated in three dimensions by defining three dimensional object types and using the generic arithmetic to operate on higher dimensional vectors.
+Objects could be combined into larger systems using procedures, if there is a way to link them together, for example by constraints.
 

### 2017-05-08 paper/main.tex

--- 2017-05-08 paper/main.tex
@@
 \section{Extensibility of the System}
-
-Possible future extensions and high level explanation of how they would be
-implemented (through minimal change to current system)
+Future extensions include collisions between objects, constraints such as joints, simulating objects in three dimensions, and being able to combine objects into larger systems of objects.
+Collisions and other kinds of constraints could be implemented by adding an interaction that produces forces that enforce these constraints.
+Objects could be simulated in three dimensions by defining three dimensional object types and using the generic arithmetic to operate on higher dimensional vectors.
+Objects could be combined into larger systems using procedures, if there is a way to link them together, for example by constraints.
 

### 2017-05-17 paper/main.tex
demo

@@
 % \newcommand{\hmwkClassInstructor}{Professor Gerry Sussman}
-\newcommand{\hmwkAuthorName}{\textbf{Author1, Author2, Author3}}
+\newcommand{\hmwkAuthorName}{\textbf{Michael Chang, Abigail Choe, Leon Shen}}
 
@@
 \maketitle
-\section{Introduction}
+% \section{Introduction}
 
-motivation: features of physics engine that incorporate ideas from class (e.g.
-generic operator, hierarchy, propagator, language for language (high--low)),
-cool visualization of concepts from class
+% motivation: features of physics engine that incorporate ideas from class (e.g.
+% generic operator, hierarchy, propagator, language for language (high--low)),
+% cool visualization of concepts from class
 
@@
 
-internal NOTE: generic procedures--we haven't actually used them in the code yet!
+% internal NOTE: generic procedures--we haven't actually used them in the code yet!
 
-NOTE: feel like this should be merged with object types and arithmetic
+% NOTE: feel like this should be merged with object types and arithmetic
 
-what we did, how we did it, why we did it
+% what we did, how we did it, why we did it
 
@@
 vector, and therefore be applicable to any $n$th dimension world
-representation. The default velocity of an object is set to 0 if it is not
+representation.
+For example, though the examples in this paper concern 2-dimensional worlds, the physics can easily be extended to 3-dimensinoal worlds.
+The default velocity of an object is set to 0 if it is not
 specified. Supporting arithmetic operations between a scalar and a vector
@@
 
-(subject to change after restructuring)
+% (subject to change after restructuring)
 
@@
 
-(TODO: add eval/apply code after restructuring)
+% (TODO: add eval/apply code after restructuring)
 

### 2017-05-18 paper/main.tex
intro

@@
 \maketitle
-% \section{Introduction}
+\section{Introduction}
+
+Physics can be viewed as the programming language of the natural world. From
+quantum mechanics to Newtonian mechanics to relativistic mechanics, the
+theories at these different levels of explanation each describe their
+primitives, means of combination, and means of abstraction over objects,
+relations, and events. Physical particles, objects, and entities are data, and
+the laws that govern how they evolve over time are procedures that operate on
+this data.
+
+Our project aims to explore this view by implementing a simple physics engine
+in scheme. We designed the physics engine to be generic and extensible. For
+example, it is very straightforward to define new worlds on the underlying
+substrate as well as to define new objects and forces within a given world.
+This paper presents the design of the infrastructure for the physics engine,
+describes how objects and forces can be easily defined and added on top of the
+infrastucture, and illustrate several demonstrations of the system on worlds
+with gravitational and electrostatic forces
 
@@
 \section{Infrastructure}
-We used the object system from pset 4 to model physical objects in our simulated world, as well as the world itself.
-Having the ability to create a hierarchy of types allows us to derive new types of physical objects to be simulated from previously defined types (see section Object Types).
+
+We used the object system from pset 4 to model physical objects in our
+simulated world, as well as the world itself. Having the ability to create a
+hierarchy of types allows us to derive new types of physical objects to be
+simulated from previously defined types (see section Object Types).
 

### 2017-05-18 paper/main.tex
punctuation

@@
 infrastucture, and illustrate several demonstrations of the system on worlds
-with gravitational and electrostatic forces
+with gravitational and electrostatic forces.
 

### 2017-05-18 paper/main.tex
General edit

@@
 % \newcommand{\hmwkIssueDate}{22 February 2017}
-\newcommand{\hmwkDueDate}{8 May 2017}
+%\newcommand{\hmwkDueDate}{8 May 2017}
 \newcommand{\hmwkClass}{6.905/6.945 Final Project Draft}
@@
     % \small{Issued\ on\ \hmwkIssueDate}\\
-    \small{Due\ on\ \hmwkDueDate}
+    %\small{Due\ on\ \hmwkDueDate}
 }
@@
 representation.
-For example, though the examples in this paper concern 2-dimensional worlds, the physics can easily be extended to 3-dimensinoal worlds.
+For example, though the examples in this paper concern 2-dimensional worlds for easy graphic display, the physics can easily be extended to 3-dimensinoal worlds and beyond.
 The default velocity of an object is set to 0 if it is not

### 2017-05-18 paper/main.tex
changed minor formatting details in report

@@
 \usepackage{array,booktabs,calc,blindtext}
-% \usepackage[shortlabels]{enumerate}
 
@@
 \newcommand{\hmwkTitle}{Extensible Physics Engine}
-% \newcommand{\hmwkIssueDate}{22 February 2017}
-%\newcommand{\hmwkDueDate}{8 May 2017}
-\newcommand{\hmwkClass}{6.905/6.945 Final Project Draft}
-% \newcommand{\hmwkClassInstructor}{Professor Gerry Sussman}
+\newcommand{\hmwkSubtitle}{Final Project Report}
+\newcommand{\hmwkClass}{6.905/6.945 - Spring 2017}
+\newcommand{\hmwkClassInstructor}{Professor Gerald Sussman}
 \newcommand{\hmwkAuthorName}{\textbf{Michael Chang, Abigail Choe, Leon Shen}}
@@
 \title{
-    \textmd{\hmwkClass:\ \hmwkTitle}\\
-    % \normalsize{\hmwkClassInstructor}\\
-    % \small{Issued\ on\ \hmwkIssueDate}\\
-    %\small{Due\ on\ \hmwkDueDate}
+    \textbf{\hmwkTitle} \\
+    \smallskip
+    \normalsize{\hmwkSubtitle} \\
+    \hmwkClass \\
+    \normalsize{\hmwkClassInstructor} \\
 }
@@
 
-% internal NOTE: generic procedures--we haven't actually used them in the code yet!
+% internal NOTE: generic procedures--we haven't actually used them in the code
+%                yet!
 
@@
 vector, and therefore be applicable to any $n$th dimension world
-representation.
-For example, though the examples in this paper concern 2-dimensional worlds for easy graphic display, the physics can easily be extended to 3-dimensinoal worlds and beyond.
-The default velocity of an object is set to 0 if it is not
-specified. Supporting arithmetic operations between a scalar and a vector
-enables our system to use a single default value 0 that is compatible with any
-vector instead of limiting our system to the dimension of the default zero
-vector specified.
+representation. For example, though the examples in this paper concern
+2-dimensional worlds for easy graphic display, the physics can easily be
+extended to 3-dimensinoal worlds and beyond. The default velocity of an object
+is set to 0 if it is not specified. Supporting arithmetic operations between a
+scalar and a vector enables our system to use a single default value 0 that is
+compatible with any vector instead of limiting our system to the dimension of
+the default zero vector specified.
 
@@
 
-There are three types of objects in the system: \texttt{thing}, \texttt{interaction}, and \texttt{world}.
-\texttt{thing} refers to all physical objects in the system.
-There is a hierarchy of physical objects that have additional properties.
-For example, type \texttt{point-charge} has an additional property \texttt{charge} and experiences electric forces while type \texttt{ball} experiences only gravity.
-Each \texttt{thing} keeps a list of \texttt{interaction} that is used to update itself.
-\texttt{interaction} refers to forces exerted on \texttt{thing}s.
-\texttt{interaction} includes a procedure that calculates the force as well as \texttt{influences}, a list of \texttt{thing}s that contribute to that force.
-Similar to \texttt{thing}, subtypes of \texttt{interaction}s form a hierarchy.
-Lastly, \texttt{world} keeps a list of all \texttt{thing}s and an \texttt{update} call to the \texttt{world} calls on each \texttt{thing} to update itself.
+There are three types of objects in the system: \texttt{thing},
+\texttt{interaction}, and \texttt{world}. \texttt{thing} refers to all physical
+objects in the system. There is a hierarchy of physical objects that have
+additional properties. For example, type \texttt{point-charge} has an
+additional property \texttt{charge} and experiences electric forces while type
+\texttt{ball} experiences only gravity. Each \texttt{thing} keeps a list of
+\texttt{interaction} that is used to update itself. \texttt{interaction} refers
+to forces exerted on \texttt{thing}s. \texttt{interaction} includes a procedure
+that calculates the force as well as \texttt{influences}, a list of
+\texttt{thing}s that contribute to that force. Similar to \texttt{thing},
+subtypes of \texttt{interaction}s form a hierarchy. Lastly, \texttt{world}
+keeps a list of all \texttt{thing}s and an \texttt{update} call to the
+\texttt{world} calls on each \texttt{thing} to update itself.
 
@@
   \centering
- \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{figs/gravity.png}
-  \caption{[\textit{left}] \texttt{earth-moon}; [\textit{center}] \texttt{binary-stars}; [\textit{right}] \texttt{solar-system}}
+  \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]
+    {figs/gravity.png}
+  \caption{[\textit{left}] \texttt{earth-moon}; [\textit{center}]
+    \texttt{binary-stars}; [\textit{right}] \texttt{solar-system}}
   \label{figure:gravity}
@@
 
-We can also combine gravity and electric charge.
-Here, we created a solar-system-like world where each of the planets repel each other and the sun attracts the planets via gravity and electric force.
-As figure \ref{figure:charges} (right) shows, this charged solar system exhibits wider and diverging orbits compared to figure \ref{figure:gravity} (right), the solar system with only gravity.
+We can also combine gravity and electric charge. Here, we created a
+    solar-system-like world where each of the planets repel each other and the
+    sun attracts the planets via gravity and electric force. As figure
+    \ref{figure:charges} (right) shows, this charged solar system exhibits
+    wider and diverging orbits compared to figure \ref{figure:gravity} (right),
+    the solar system with only gravity.
 
@@
   \centering
- \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{figs/charges.png}
-  \caption{[\textit{left}] \texttt{charges-1}; [\textit{center}] \texttt{charges-2}; [\textit{right}] \texttt{charged-solar-system}}
+  \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]
+    {figs/charges.png}
+  \caption{[\textit{left}] \texttt{charges-1};
+    [\textit{center}] \texttt{charges-2};
+    [\textit{right}] \texttt{charged-solar-system}}
   \label{figure:charges}
@@
 \section{Extensibility of the System}
-Future extensions include collisions between objects, constraints such as joints, simulating objects in three dimensions, and being able to combine objects into larger systems of objects.
-Collisions and other kinds of constraints could be implemented by adding an interaction that produces forces that enforce these constraints.
-Objects could be simulated in three dimensions by defining three dimensional object types and using the generic arithmetic to operate on higher dimensional vectors.
-Objects could be combined into larger systems using procedures, if there is a way to link them together, for example by constraints.
+Future extensions include collisions between objects, constraints such as
+joints, simulating objects in three dimensions, and being able to combine
+objects into larger systems of objects. Collisions and other kinds of
+constraints could be implemented by adding an interaction that produces forces
+that enforce these constraints. Objects could be simulated in three dimensions
+by defining three dimensional object types and using the generic arithmetic to
+operate on higher dimensional vectors. Objects could be combined into larger
+systems using procedures, if there is a way to link them together, for example
+by constraints.
 

### 2017-05-19 paper/main.tex
tweaked report slightly

@@
 example, it is very straightforward to define new worlds on the underlying
-substrate as well as to define new objects and forces within a given world.
-This paper presents the design of the infrastructure for the physics engine,
+substrate as well as to define new objects and forces within a given world. We
+attempt to impose as little constraint on the behavior of the objects as
+possible, by allowing them to control how they are updated over time.  This
+paper presents the design of the infrastructure for the physics engine,
 describes how objects and forces can be easily defined and added on top of the
@@
 with gravitational and electrostatic forces.
+
+All code for this project can be viewed at
+\texttt{https://github.com/mbchang/6.945-physics}.
 
@@
 types of object. Moreover, in the second model, the engine does not need to be
-notified and updated everytime a new object is added. Instead, update or
+notified and updated every time a new object is added. Instead, update or
 addition of new \texttt{interaction}s is taken care of at the object level when
@@
 
+Seeing as objects can choose how to update themselves, objects which behave in
+exotic ways, not necessarily obeying any physical laws or principles, could be
+added and integrated with physical objects.
+
 \end{document}
